---
title: "Homework 1"
author: "Yufei Lin, Jingfeng Xia"
date: "Nov 7 2020"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
library(pander)
library(knitr)
library(skimr)
library(kableExtra)
library(tinytex)
library(dplyr)
library(purrr)
local({
  hook_inline = knitr::knit_hooks$get('inline')
  knitr::knit_hooks$set(inline = function(x) {
    res = hook_inline(x)
    if (is.numeric(x)) sprintf('$%s$', res) else res
  })
})
```

```{r libraries}
#install.packages('pls')
#install.packages('randomForest')
library(pls)
library(randomForest)
library(gam)
```

## Data Processing

### Read in Data

```{r readData}
curDir = getwd()
setwd(curDir)
vault = read.csv("./SourceData/test.csv")
HousePricing = read.csv("./SourceData/train.csv")
```

### Seperate into Test and Training Set

```{r separateData}
set.seed(1)
randS = sample(1:nrow(HousePricing), nrow(HousePricing)/2)
train = HousePricing[randS,]
test = HousePricing[-randS,]
summary(test)
```

### Feature Engineering

In this section, we convert all unnecessary NAs into suitable value based on the observation from data description on Kaggle website. 

```{r fe}
# Currently I have solved this by replacing all NA with 0 in excel,
# for some reason the replacement does not work in R. 
```


## Prediction Algorithms

### 1. PCA

```{r PCA}
set.seed(2)
# Need help with PCA analysis
#pcr.fit=pcr(SalePrice~., data=HousePricing, scale=TRUE, validation ="CV")
```

### 2. Random Forest

```{r rf}
rfTrain=randomForest(SalePrice~.,data=train, mtry=6,importance =TRUE, na.action=na.roughfix)
rfYhat = predict(rfTrain ,newdata=test)
print("The test MSE is shown in the following: ")
mean((rfYhat-test$SalePrice)^2)
```

```{r rf2}
temp = importance(rfTrain)
temp = as.data.frame(temp)
names(temp)[names(temp)=="%IncMSE"] <- "importance"
sort(temp$importance, decreasing = TRUE)
```
```{r rf3}
pander(temp, title="Importance of each factor according to random forest")
```

From the random forest analysis, we have discovered that the top three most important factors for predicting sale price are the following:

\begin{enumerate}
\item GrLivArea
\item Neighbourhood
\item OverallQual
\end{enumerate}

Therefore, we will make GAM models according to these three factors. 

```{r GAM}
fit1 = gam(SalePrice ~ GrLivArea + Neighborhood + OverallQual, data = HousePricing)
print("Deviance of Model 1")
deviance(fit1)
```

From this we know that the deviance is quite large, we need a better model. 

## Cross Validation


































































